---
title: "Survival Analysis of Post-Myocardial Infarction Patients"
author: "Alvein, Orr, Pham"
date: "5/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Abstract
** Two sentences each, at most!

[Brief Intro]

[Objective + methods]

Using survival and hazard analysis tools, we see to better understand the nature of survival for patients that have survived the infarction. As rate of infarction cases increase, so will the rate of patients that will require monitoring.We found that 

[Results]
We have found that...

[Conclusion]
It's clear that survival rates is not the same after a heart attack.??


# Introduction

Myocardial infarctions are becoming largely common among U.S. populations. The number of myocardial infarctions are remarkedly increased from [start date] to [end date] by [x] amount (add citation). In 2015, approximately 23% of all fatalities in the United States was related to some degree of heart disease (cdc. cite please). Unsurprisingly, clinical studies have shown harmful symptoms in post-infarction survival patients. Reports of decreased blood pressure efficiency and abnormal heart rate have raised concerns of variable-dependent survivability among post-infarction patients. Our obtained dataset to examine the tangible difference in survability rates from the course of year following an infarction episode.

[Insert more kaggle context]

```{r cdc.deathsummary, echo=FALSE}


summary(cars)
```

We have obtained a data set examining the the length and survival of post-myocardial patients over the course of one year.

## Objective
We seek to apply survival analysis techniques to this given dats set to examine the characteristics of survival rates among post-infarction patients. Additionally, we hope to achieve better understanding of survival rates as stratified among several classifications.

## Dataset
We have obtained our data set from Kaggle. The data set contains # of observations and records 13 variables. Please refer to the data set below for more details.

[add table summarizing the headers]

```{r Dataset.summary, echo=FALSE}
summary(cars)
```

[add the data set??]

```{r Dataset.actual, echo=FALSE}
summary(cars)
```

[add details from kaggle]

## Mathematical Definitions
### Censoring
### Cruves
### equations

[latex]

# Methodology
## Summary statistics
## Kaplan-Meier
## Weibull Fits
## Cox PH

# Results
## Table of Summary Statistics

```{r stats.summary, echo=FALSE}
summary(cars)
```

## KM Curve

```{r km.plot, echo=FALSE}
summary(cars)
```

## Weibull Curve

```{r wb.plot, echo=FALSE}
summary(cars)
```

## Cox Proportional Hazard

```{r coxph.plot, echo=FALSE}
summary(cars)
```

## Model Diagnostics
### AIC, BIC, and Confidence Intervals

```{r diagnostic.table, echo=FALSE}
summary(cars)
```

### Residual Analysis/QQ Plot

```{r diagnostic.resid, echo=FALSE}
summary(cars)
```

```{r diagnostic.qqplot, echo=FALSE}
summary(cars)
```

# Discussion

# References