---
title: "Survival Analysis of Post-Myocardial Infarction Patients"
author: "Alvein, Orr, Pham"
date: "5/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(knitr)
library(tidyverse)
library(dplyr)
library(kableExtra)
library(dplyr)
```
# Abstract

## Background
The rates of myocardial infarction is becoming an increasing common occurence in the United States. As medical knowledge and techniques improve to meet need of infarction episode, so does the need to understand the survivbility of patients who have survived such episodes.

## Objectives
Our goal is to provide detailed survival statistics of post-myocardial infarction patients as well as provide an accurate regression model to best prediction of survival outcomes of a single year following an infarction episode.

## Methods
Data from 133 post-myocardial infarction patients measure the time in months until death in a one year monitoring period of follow-up. We use a combination of nonparametric (Kaplan-Meier) and parametric methods (Weibull/Cox PH) to determine estimates of survival among age, gender, and myocardial strata (contraction depth, muscular activity, ). We consider a slew of statistical and graphical results before determining the most appropriate method of modeling.

## Results
Out of all of our methods, we have determined that [] is the most appropriate model for prediction of patient survival. We have AIC values of. We have BIC values. Thus, this model is the best.


## Conclusion
[summary statistics]
[review of our model + specific surval rates]




# Introduction

Myocardial infarctions are becoming largely common among U.S. populations. The number of myocardial infarctions are remarkedly increased from [start date] to [end date] by [x] amount (add citation). In 2015, approximately 23% of all fatalities in the United States was related to some degree of heart disease (cdc. cite please). Unsurprisingly, clinical studies have shown harmful symptoms in post-infarction survival patients. Our obtained dataset to examine the tangible difference in survability rates from the course of year following an infarction episode.

By applying survival analysis techniques to this data set, we seek to achieve improved understanding of the characteristics exhibited by patients in a one year post-infarction interval.We also propose a model to better predict the probability of a survival of patients based on these variable characteristics.


## Dataset
We have obtained our data set from Kaggle. The data set contains 131 of observations and records 8 variables. It should be noted that two patients data were removed as there was no recorded months survival date.

Below is a table summary of each table header variable label and a definition of that variable.

```{r Dataset.summary, echo=FALSE}
df.sum <- data.frame(read_excel("df.sum.xlsx"))
kable(df.sum, caption = "Dataset Variable Labels")
```

Since the time of infartion varies, some patients were followed for less than a year. This provides a clear censoring and truncation provide. We will address this concern in detail in the later in this section. It should be noted that we have a slew of missing values. Since a single patient has shown as missing, we have opted to impute the values for this data row. With this in mind, our predictive and summary models will have less than ideal accuracy.

The full dataset is available in the appendix of this paper.


# Methodology

## Censoring
Left Censored: Because all the patients had their heart attacks at different times, it is possible that some patients have survived less than one year but they are still alive. Check the second variable to confirm this. Such patients cannot be used for the prediction task mentioned above.

## Imputation
There are several missing values in the data set. The variables we chose to more 
## Summary statistics
## Kaplan-Meier
We conduct nonparametric Kaplan-Meier fit on our data over multiple strata. We first conduct a fit over all patients in regards to censoring. Then we work on gender groups, age, and general ventriculular condition.

## Weibull Fits
## Cox PH


# Results
## Table of Summary Statistics

```{r stats.summary, echo=FALSE}
summary(cars)
```

## KM Curve

KM Overall
```{r km.all, echo=FALSE}
summary(cars)
```

KM Age
```{r km.age, echo=FALSE}
summary(cars)
```

KM Gender
```{r km.gender, echo=FALSE}
summary(cars)
```

KM Ventricular Activity
```{r km.plot, echo=FALSE}
summary(cars)
```


## Weibull Curve

```{r wb.plot, echo=FALSE}
summary(cars)
```

## Cox Proportional Hazard

```{r coxph.plot, echo=FALSE}
summary(cars)
```

## Model Diagnostics
### AIC, BIC, and Confidence Intervals

```{r diagnostic.table, echo=FALSE}
summary(cars)
```

### Residual Analysis/QQ Plot

```{r diagnostic.resid, echo=FALSE}
summary(cars)
```

```{r diagnostic.qqplot, echo=FALSE}
summary(cars)
```

# Discussion

# References
[https://www.cdc.gov/heartdisease/facts.htm]

Heron, M. Deaths: Leading causes for 2017 pdf icon[PDF – 3 M]. National Vital Statistics Reports;68(6). Accessed November 19, 2019.

Salzberg, S. (1988). Exemplar-based learning: Theory and implementation (Technical Report TR-10-88). Harvard University, Center for Research in Computing Technology, Aiken Computation Laboratory (33 Oxford Street; Cambridge, MA 02138).

Kan, G., Visser, C., Kooler, J., & Dunning, A. (1986). Short and long term predictive value of wall motion score in acute myocardial infarction. British Heart Journal, 56, 422-427.

Fryar CD, Chen T-C, Li X. Prevalence of uncontrolled risk factors for cardiovascular disease: United States, 1999–2010 pdf icon[PDF-494K]. NCHS data brief, no. 103. Hyattsville, MD: National Center for Health Statistics; 2012. Accessed May 9, 2019.


# Appendix

```{r Dataset.actual, echo=FALSE, results='asis'}
df <- data.frame(read_excel("df.xlsx"))
kable(df, caption = "Dataset")
```
